Understanding the impact of electronically-delivered parenting advice

### 2019-03-09 paper/cogsci2019/cogsci2019.tex

Young children spend a large portion of their waking time at play,
variously manipulating objects, exploring their environment, and
interacting with caregivers and peers. Playing with objects allows them
to discover hidden object properties and relations, and to build a
causal understanding of how objects interact (e.g., Schulz \& Bonawitz,
2007). Meanwhile, play also gives children an opportunity to set and
achieve goals (e.g., build a tower) and to practice a wide range of
motor skills (e.g., stacking) that will help them navigate the world
(Singer, Golinkoff, \& Hirsh-Pasek, 2006). Social play can help children
learn about human relationships, both through imitation of adult
behaviors and by experiencing and learning to process emotional events
such as failures (Singer et al., 2006). Of course, young children are
rarely playing in isolation: caregivers often provide encouragement and
guidance while scaffolding a child's play (Kaye, 1970; Wood, Bruner, \&
Ross, 1976). Indeed, parenting practices early in childhood have been
shown to play an important role in shaping the future outcomes of young
children (Hart \& Risley, 1995; Heckman, 2006). While interventions
often have trouble reaching many parents of very young children, the
proliferation of mobile devices offers a good avenut for the digital
delivery of parenting advice (Breitenstein, Gross, \& Christophersen,
2014).

@@

The video of each session was manually coded for episodes of joint
attention using the Datavyu software (Team, 2014). The video taken at
floor level was coded by default, but the other video was referred to if
the participants were not visible or if there was technical difficulty
with the first camera. Each session's video was coded for episodes of
coordinated joint attention, episodes of passive joint attention, and
parental bids for joint attention. Parental bids for joint attention
were defined as any attempt to initiate joint attention (i.e labeling,
pointing, or otherwise drawing attention to an object) that did not
result in passive or coordinated joint attention. If more than 3 seconds
elapsed between bids, they were coded as separate attempts. An episode
of joint attention was considered passive if both participants visually
focused on an object for a minimum of 3 seconds but the child did not
acknowledge the parent. If either participant looked away from the
object for less than 3 seconds and then returned to the same object it
was considered part of the same period of joint attention. Episodes
joint attention were considered coordinated if both participants
visually focused on an object for a minimum of 3 seconds and at some
point in the interaction the child indicated awareness of interaction
with some overt behavior toward the parent. This could be looks to the
parent's face, gestures, vocalizations, or turn-taking. If either
participant looked away for less than 3 seconds and then returned to the
same object, it was coded as part of the same period of joint attention.
A second coder independently coded a third of the videos (i.e., 20 of
the 60 videos, approximately equally distributed across ages) to
establish reliability. The two coders had a reliability of ICC = 0.80
with 95\% confident interval (CI) = {[}0.57,0.92{]} (p \textless{} 0.05)
for number of parent bids for JA; ICC = 0.20 with 95\% CI =
{[}-0.26,0.58{]} for number of passive JA episodes; ICC = 0.66 with 95\%
CI = {[}0.32,0.85{]} (p \textless{} 0.05) for number of coordinated JA
episodes; ICC = 0.24 with 95\% CI = {[}-0.21,0.61{]} for total duration
of passive JA episodes, and ICC = 0.62 with 95\% CI = {[}0.27,0.83{]} (p
\textless{} 0.05) for total duration of coordinated JA episodes.

>>>>

Young children spend a large portion of their waking time at play,
variously manipulating objects, exploring their environment, and
interacting with caregivers and peers. Playing with objects allows them
to discover hidden object properties and relations, and to build a
causal understanding of how objects interact (e.g., Schulz \& Bonawitz,
2007). Meanwhile, play also gives children an opportunity to set and
achieve goals (e.g., build a tower) and to practice a wide range of
motor skills (e.g., stacking) that will help them navigate the world
(Singer, Golinkoff, \& Hirsh-Pasek, 2006). Social play can help children
learn about human relationships, both through imitation of adult
behaviors and by experiencing and learning to process emotional events
such as failures (Singer et al., 2006). Of course, young children are
rarely playing in isolation: caregivers often provide encouragement and
guidance while scaffolding a child's play (Kaye, 1970; Wood, Bruner, \&
Ross, 1976). Indeed, parenting practices early in childhood have been
shown to play an important role in shaping the future outcomes of young
children (Hart \& Risley, 1995; Heckman, 2006). While interventions
often have trouble reaching many parents of very young children, the
proliferation of mobile devices offers a good avenut for the digital
delivery of parenting advice (Breitenstein, Gross, \& Christophersen,
2014). However, the efficacy of digital parenting advice has not been
widely proven.

@@

The video of each session was manually coded for episodes of joint
attention using the Datavyu software (Team, 2014). The video taken at
floor level was coded by default, but the other video was referred to if
the participants were not visible or if there was technical difficulty
with the first camera. Each session's video was coded for episodes of
coordinated joint attention, episodes of passive joint attention, and
parental bids for joint attention. Parental bids for joint attention
were defined as any attempt to initiate joint attention (i.e labeling,
pointing, or otherwise drawing attention to an object) that did not
result in passive or coordinated joint attention. If more than 3 seconds
elapsed between bids, they were coded as separate attempts. An episode
of joint attention was considered passive if both participants visually
focused on an object for a minimum of 3 seconds but the child did not
acknowledge the parent. If either participant looked away from the
object for less than 3 seconds and then returned to the same object it
was considered part of the same period of joint attention. A joint
attention episode was considered coordinated if both participants
visually focused on an object for a minimum of 3 seconds and at some
point in the interaction the child indicated awareness of interaction
with some overt behavior toward the parent such as looking at their
face, gesturing, vocalizing, or turn-taking.

A second coder independently coded a third of the videos (i.e., 20 of
the 60 videos, approximately equally distributed across ages) to
establish reliability. The two coders had a reliability of ICC = 0.80
with 95\% confident interval (CI) = {[}0.57,0.92{]} (p \textless{} 0.05)
for number of parent bids for JA; ICC = 0.20 with 95\% CI =
{[}-0.26,0.58{]} for number of passive JA episodes; ICC = 0.66 with 95\%
CI = {[}0.32,0.85{]} (p \textless{} 0.05) for number of coordinated JA
episodes; ICC = 0.24 with 95\% CI = {[}-0.21,0.61{]} for total duration
of passive JA episodes, and ICC = 0.62 with 95\% CI = {[}0.27,0.83{]} (p
\textless{} 0.05) for total duration of coordinated JA episodes.
